Score-based tests of measurement invariance: Use in practice

### 2013-08-27 frontiers/frontiers.Rnw
Try a small simulation in simall-ordinal.R, but when parameter is loading or var, the plots seems wrong. I haven't figured out why.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@243 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%% TODO focus on ordinal
Multiple test statistics are available, depending on how one
summarizes the behavior of the cumulative sum of scores.  For example, one could
take the absolute maximum that the cumulative sum attains for any
parameter of interest, resulting in a {\em double max} statistic (the maximum is taken across parameters and individuals).
Alternatively, one could sum the (squared) cumulative sum across parameters
of interest and take the maximum or the average across individuals, resulting in a {\em maximum Lagrange multiplier} statistic and Cram\'{e}r-von Mises statistic, respectively \cite<see>[for further discussion]{MerZei13}.  These statistics are given by
\begin{eqnarray}
    \label{eq:dmax}
    \mathit{DM} & = & \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} \\
        \label{eq:cvm}
    \mathit{CvM}     & = & n^{-1} \sum_{i = 1,\dots, n} \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2, \\
    \label{eq:maxlm}
    \max \mathit{LM} & = & \max_{i = \underline{i}, \dots, \overline{\imath}} ~
      \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
      \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2.
\end{eqnarray}

@@

We focus on measurement invariance of the factor loadings associated with the GQ-6 scale, using a one-factor model.
While the ``age group'' variable against which we are testing measurement invariance is best considered ordinal, for demonstration we consider its treatment as categorical, continuous, and ordinal.  Each of these treatments is described below in a separate section.

@@

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.
<<critvals,echo=FALSE>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo",
                 fnl=opval1)
@

@@

%% gefp1<-gefp(reduced,fit=NULL,order.by=as.numeric(yg$agegroup),
%%        vcov=info_full,sandwich=FALSE,parm=1:4)
%% pval.ord=rep(NA,2)
%% pval.ord[1]<-sctest(gefp1,functional=ordwmax(gefp1))$p.value

%% pval.ord[2]<-sctest(gefp1, functional=opval1)$p.value
%% pval.ord
%% @

@@

In Simulation 1, we examined the extent

>>>>

%% TODO focus on ordinal
Multiple test statistics are available, depending on how one
summarizes the behavior of the cumulative sum of scores.  For example, 
one could take the absolute maximum that the cumulative sum attains for any
parameter of interest, resulting in a {\em double max} statistic 
(the maximum is taken across parameters and individuals).
Alternatively, one could sum the (squared) cumulative sum across parameters
of interest and take the maximum or the average across individuals, resulting in a 
{\em maximum Lagrange multiplier} statistic and Cram\'{e}r-von Mises statistic, respectively 
\cite<see>[for further discussion]{MerZei13}.  These statistics are given by

\begin{eqnarray}
    \label{eq:dmax}
    \mathit{DM} & = & \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} \\
        \label{eq:cvm}
    \mathit{CvM}     & = & n^{-1} \sum_{i = 1,\dots, n} \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2, \\
    \label{eq:maxlm}
    \max \mathit{LM} & = & \max_{i = \underline{i}, \dots, \overline{\imath}} ~
      \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
      \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2.
\end{eqnarray}

@@

We focus on measurement invariance of the factor loadings associated with t
he GQ-6 scale, using a one-factor model.While the ``age group'' variable against 
which we are testing measurement invariance is best considered ordinal, 
for demonstration we consider its treatment as categorical, continuous, and ordinal.  
Each of these treatments is described below in a separate section.

@@

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.

<<critvals,echo=FALSE>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo",
                 fnl=opval1)
@

@@

%% gefp1<gefp(reduced,fit=NULL,order.by=as.numeric(yg$agegroup),
%%        vcov=info_full,sandwich=FALSE,parm=1:4)
%% pval.ord=rep(NA,2)
%% pval.ord[1]<-sctest(gefp1,functional=ordwmax(gefp1))$p.value

%%pval.ord[2]<-sctest(gefp1, functional=opval1)$p.value
%% pval.ord
%% @

@@
